Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Proof-of-Concept V}}

### 2014-05-03 Paper.tex
PoC-5 cleanups.

When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c. On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.

@@

Thus we may define a world-state collapse function $L_S$:
\begin{equation}
L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
\end{equation}
where
\begin{equation}
p(a) \equiv \begin{cases}
\big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b) \big) \big) & \text{if} \; \boldsymbol{\sigma}[a]_s = \varnothing\\
\big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big) & \text{otherwise}
\end{cases}
\end{equation}
This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
\begin{equation}
\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
\end{equation}

Where $v$ is the account validity function:
\begin{eqnarray}
\quad v(x) & \equiv & x_n \in \mathbb{P} \wedge x_b \in \mathbb{P} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
\end{eqnarray}

@@

A transaction (formally, $T$) is a single cryptographically signed instruction sent by an actor external to Ethereum. An external actor can be a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server. There are two types of transactions: those which result in message calls and those which result in the creation of new contracts. Both types specify a number of common fields:

@@

Whereas a message call transaction contains:

@@

Appendix \ref{app:signing} specifies the function mapping transactions to the sender $S$, which happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign. For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.

@@

Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{b}$, $T_\mathbf{i}$ and $T_\mathbf{d}$ and the address hash $T_\mathbf{t}$:
\begin{equation}
\begin{array}[t]{lclclc}
T_n \in \mathbb{P} & \wedge & T_v \in \mathbb{P} & \wedge & T_p \in \mathbb{P} & \wedge \\
T_g \in \mathbb{P} & \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge \\
T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B} & \wedge \\
T_t \in \mathbb{B}_{20} &&&&\\
\end{array}
\end{equation}

@@

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with a set of transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}). The block header contains several pieces of information:

@@

The other two components in the block are simply a list of uncle block headers (of the same format as above) and a list of the transaction information (as defined previously). Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
\end{equation}

We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
\begin{equation}
\begin{array}[t]{lclc}
H_u &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert: (i, B_\mathbf{T}[i])\}) & \wedge \\
H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) &
\end{array}
\end{equation}

where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.

% TODO: formal.

Furthermore:

\begin{equation}
\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
\end{equation}

@@

TODO: Transaction manifest more than just transactions: $(T[i], s[i], g[i])$

@@

The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively. Much like the transaction preparation function $L_T$, we assert the types and order of the structure for when the RLP transformation is required:
\begin{eqnarray}
\quad L_H(H) & \equiv & (H_p, H_u, H_b, H_r, H_t, H_d, H_s, H_x, H_n) \\
\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
\end{eqnarray}

@@

The validity is determined as:
\begin{equation}
\begin{array}[t]{rcl}
S(T) & \neq & \varnothing \quad \wedge \\
\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
g_0 & \leq & T_g \quad \wedge \\ 
v_0 & \leq & \boldsymbol{\sigma}[S(T)]_b
\end{array}
\end{equation}

The sum of the transaction gas limit and the gas utilised in this block prior must be no greater than the block's \textbf{gasLimit}, $H_l$.

% TODO: specify formally.

@@

Code execution depletes gas; thus it may exit before the code has come to a natural halting state. In this exceptional case we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation. The gas remaining should be zero. If the creation was conducted as the manifestation of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.

@@

The given \textbf{gasUsed} and \textbf{minGasLimit} must correspond faithfully to the transactions listed. In the case of ${B_H}_m$, the minimum gas limit, all transactions included in the transaction manifest must have a gas price, ${T_0}_p$, that is at least this value:
\begin{equation}
\forall_{T \in B_\mathbf{T}} {T_0}_p \geq {B_H}_m
\end{equation}

In the case of ${B_H}_u$, the total gas used in the block, it must be equal to the accumulated gas used according to the final transaction:
\begin{equation}
{B_H}_u = \ell(\mathbf{T})_1
\end{equation}

@@

As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined. We begin the definition by supposing we have a series of transactions $\mathbf{T}$.

\begin{equation}
\text{Let} \quad \mathbf{T} = (\mathbf{T}[0], \mathbf{T}[1], ...)
\end{equation}

These are the transactions that comprise the block, thus:
\begin{eqnarray}
\quad \quad B & \equiv & \Big(H, \big( \mathbf{M}[0], \mathbf{M}[1], ..., \mathbf{M}[\lVert \mathbf{T}\rVert - 1]\big), \mathbf{U}\Big) \\
\quad \quad \mathbf{M}[n] & \equiv & (\mathbf{T}[n], \mathbf{s}[n], \mathbf{g}[n])
\end{eqnarray}

Here, $\mathbf{s}$ and $\mathbf{g}$ are the corresponding sequences of states and cumulative gas used for each transaction. $\mathbf{s}$ is defined simply as the state resulting from applying the corrresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
\begin{equation}
\mathbf{s}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{s}[n - 1], \mathbf{T}[n]) & \text{otherwise} \end{cases}
\end{equation}

In the case of $\mathbf{g}$, we take a similar approach defining each item as the gas used in evaluatng the corresponding transaction summed with the previous item (or zero, if it's the first), giving us a running total:
\begin{equation}
\mathbf{g}[n] = \begin{cases} 0 & \text{if} \quad n < 0 \\ \Upsilon^g(\mathbf{s}[n - 1], \mathbf{T}[n]) + \mathbf{g}[n-1] & \text{otherwise} \end{cases}
\end{equation}

Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{T})_1$:
\begin{equation}
\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(B_\mathbf{T})_1)
\end{equation}

@@

Rather than being hard-specified in the protocol, the block limit will become based upon some fixed ratio about a long-term moving average of actual gas usage in blocks, allowing it to slowly increase according to the network's ability to manage the transaction volume.

@@

\begin{eqnarray}
C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) & \equiv & G_{memory}(\boldsymbol{\mu}'_i - \boldsymbol{\mu}_i) + \begin{cases}
C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
G_{call} & \text{if} \quad w = \text{\small CALL}\\
G_{create} & \text{if} \quad w = \text{\small CREATE}\\
G_{sha3} & \text{if} \quad w = \text{\small SHA3}\\
G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
G_{stop} & \text{if} \quad w = \text{\small STOP}\\
G_{suicide} & \text{if} \quad w = \text{\small SUICIDE}\\
G_{step} & \text{otherwise}
\end{cases} \\
w & \equiv & \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{eqnarray}

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{f0s: System operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0xf0 & {\small CREATE} & 3 & 1 & Create a new contract. \\
&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g) \equiv \Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, \mathbf{b})$ \\
&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
&&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
&&&& $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created contract, otherwise. \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2]) \div 32 })$ \\
\midrule
0xf1 & {\small CALL} & 7 & 1 & Message-call into contract. \\
&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
&&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
&&&& Let $g_k = \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0 \\ \boldsymbol{\mu}_g - \min(\boldsymbol{\mu}_g, \boldsymbol{\mu}_\mathbf{s}[2]) & \text{otherwise} \end{cases}$ \\
&&&& $(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[0], g, I_p, \boldsymbol{\mu}_g - g_k, \boldsymbol{\mu}_\mathbf{s}[1], \mathbf{i})$ \\
&&&& $\boldsymbol{\mu}'_g \equiv g' + g_k$ \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
&&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
&&&& $x=1$ otherwise. \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ \max(\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6]) \div 32 })$ \\
\midrule
0xf2 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
&&&& This has the effect of halting the execution at this point with output defined.\\
&&&& See section \ref{ch:model}. \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \div 32 })$ \\
\midrule
0xff & {\small SUICIDE} & 1 & 0 & Halt execution and obliterate account. \\
&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0]]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b + \boldsymbol{\sigma}[I_a]_b$ \\
&&&& $\boldsymbol{\sigma}'[I_a] \equiv \varnothing$ \\
\bottomrule
\end{tabular*}

@@

\end{document}

>>>>

When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c. On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.

@@

Thus we may define a world-state collapse function $L_S$:
\begin{equation}
L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
\end{equation}
where
\begin{equation}
p(a) \equiv  \big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
\end{equation}

This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
\begin{equation}
\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
\end{equation}
where $v$ is the account validity function:
\begin{eqnarray}
\quad v(x) & \equiv & x_n \in \mathbb{P} \wedge x_b \in \mathbb{P} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
\end{eqnarray}

@@

A transaction (formally, $T$) is a single cryptographically-signed instruction sent by an actor external to Ethereum. An external actor can be a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server. There are two types of transactions: those which result in message calls and those which result in the creation of new contracts. Both types specify a number of common fields:

@@

However, a message call transaction contains:

@@

Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign. For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.

@@

Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{b}$, $T_\mathbf{i}$ and $T_\mathbf{d}$ and the address hash $T_\mathbf{t}$:
\begin{equation}
\begin{array}[t]{lclclc}
T_n \in \mathbb{P} & \wedge & T_v \in \mathbb{P} & \wedge & T_p \in \mathbb{P} & \wedge \\
T_g \in \mathbb{P} & \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge \\
T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B} & \wedge \\
T_t \in \mathbb{B}_{20} &&&&
\end{array}
\end{equation}

@@

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{R}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}). The block header contains several pieces of information:

@@

The other two components in the block are simply a list of uncle block headers (of the same format as above) and a series of the transaction \textit{receipts}. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_H, B_\mathbf{R}, B_\mathbf{U})
\end{equation}

\subsubsection{Transaction Manifest}

The transaction receipt is a tuple of three items comprising the transaction, $R_T$, together with the post-transaction state, $R_{\boldsymbol{\sigma}}$, and the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$:
\begin{equation}
R \equiv (R_T, R_{\boldsymbol{\sigma}}, R_u)
\end{equation}

The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
\begin{equation}
L_R(R) \equiv (L_T(R_T), \mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u)
\end{equation}
thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the second item.

We assert $R_u$, the cumulative gas used is a positive integer:
\begin{equation}
R_u \in \mathbb{P}
\end{equation}

For convenience, I also name $B_\mathbf{T}$ the series of transactions encoded by the transaction receipts $B_\mathbf{R}$:
\begin{equation}
B_\mathbf{T} \equiv (B_\mathbf{R}[0]_T, B_\mathbf{R}[1]_T, ...)
\end{equation}

Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.

\subsubsection{Holistic Validity}

We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
\begin{equation}
\begin{array}[t]{lclc}
H_u &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert: (i, B_\mathbf{R}[i])\}) & \wedge \\
H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) &
\end{array}
\end{equation}

Furthermore:
\begin{equation}
\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
\end{equation}

@@

@@

\subsubsection{Serialisation}

The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively. Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
\begin{eqnarray}
\quad L_H(H) & \equiv & (H_p, H_u, H_b, H_r, H_t, H_d, H_s, H_x, H_n) \\
\quad L_B(B) & \equiv & \big( L_H(B_H), L_R^*(B_\mathbf{R}), L_H^*(B_\mathbf{U}) \big)
\end{eqnarray}

@@

The validity is determined as:
\begin{equation}
\begin{array}[t]{rcl}
S(T) & \neq & \varnothing \quad \wedge \\
\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
g_0 & \leq & T_g \quad \wedge \\ 
v_0 & \leq & \boldsymbol{\sigma}[S(T)]_b \\
T_l \leq {B_H}_l - \ell(B_\mathbf{T})_2
\end{array}
\end{equation}

Note the final condition; the sum of the transaction's gas limit, $T_l$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{T})_2$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$:

@@

Code execution depletes gas; thus it may exit before the code has come to a natural halting state. In this exceptional case we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation. The gas remaining should be zero. If the creation was conducted as the receiptation of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.

@@

%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.

The given \textbf{gasUsed} and \textbf{minGasLimit} must correspond faithfully to the transactions listed. In the case of ${B_H}_m$, the minimum gas limit, all transactions included in the transaction receipt must have a gas price, ${T_0}_p$, that is at least this value:
\begin{equation}
\forall_{T \in B_\mathbf{T}} T_p \geq {B_H}_m
\end{equation}

In the case of ${B_H}_u$, the total gas used in the block, it must be equal to the accumulated gas used according to the final transaction:
\begin{equation}
{B_H}_u = \ell(\mathbf{R})_u
\end{equation}

@@

As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.

As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states and cumulative gas used after each transaction. The former is defined simply as the state resulting from applying the corrresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
\begin{equation}
B_\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
\end{equation}

In the case of $\mathbf{g}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
\begin{equation}
B_\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
\begin{array}[b]{l}
\Upsilon^g(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + B_\mathbf{R}[n-1]_u
\end{array}
 & \text{otherwise} \end{cases}
\end{equation}

Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{T})_1$:
\begin{equation}
\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(B_\mathbf{T})_{\boldsymbol{\sigma}})
\end{equation}

@@

@@

\begin{eqnarray}
C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) & \equiv & G_{memory}(\boldsymbol{\mu}'_i - \boldsymbol{\mu}_i) + \begin{cases}
C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad w = \text{\small CALL}\\
G_{create} & \text{if} \quad w = \text{\small CREATE}\\
G_{sha3} & \text{if} \quad w = \text{\small SHA3}\\
G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
G_{stop} & \text{if} \quad w = \text{\small STOP}\\
G_{suicide} & \text{if} \quad w = \text{\small SUICIDE}\\
G_{step} & \text{otherwise}
\end{cases} \\
w & \equiv & \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{eqnarray}

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{f0s: System operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0xf0 & {\small CREATE} & 3 & 1 & Create a new contract. \\
&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g) \equiv \Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, \mathbf{b})$ \\
&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
&&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
&&&& $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created contract, otherwise. \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2]) \div 32 })$ \\
\midrule
0xf1 & {\small CALL} & 7 & 1 & Message-call into contract. \\
&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
&&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
&&&& $(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i})$ \\
&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
&&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
&&&& $x=1$ otherwise. \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ \max(\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6]) \div 32 })$ \\
\midrule
0xf2 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
&&&& This has the effect of halting the execution at this point with output defined.\\
&&&& See section \ref{ch:model}. \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \div 32 })$ \\
\midrule
0xff & {\small SUICIDE} & 1 & 0 & Halt execution and obliterate account. \\
&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0]]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b + \boldsymbol{\sigma}[I_a]_b$ \\
&&&& $\boldsymbol{\sigma}'[I_a] \equiv \varnothing$ \\
\bottomrule
\end{tabular*}

@@

\section{Javascript API}\label{app:jsapi}

Yet TODO.
\end{document}
